---
title: "Team Project"
team members: Jiaqi Ruan, Ke Wu, Zhengqi Zhang
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    number_sections: yes
    toc: yes
  word_document:
    toc: yes
---

# Intro
```{r}
rm(list = ls())     # clear the workspace 
library(ISLR)       
library(tidyverse)
library(ggplot2)
```

# Data Preparation
Examine the dataset. It contains four variables, `LIMIT_BAL`, `SEX`,`EDUCATION`, and `MARRIAGE`, etc. 
```{r}
data <- read.csv("UCI_Credit_Card.csv")
credit = subset(data, select = c("LIMIT_BAL", "SEX", "EDUCATION", "MARRIAGE", "AGE", "PAY_0",
                                 "PAY_2", "PAY_3", "PAY_4", "PAY_5", "PAY_6",
                                 "BILL_AMT1", "BILL_AMT2", "BILL_AMT3",
                                 "BILL_AMT4", "BILL_AMT5", "BILL_AMT6",
                                 "PAY_AMT1", "PAY_AMT2", "PAY_AMT3","PAY_AMT4",
                                 "PAY_AMT5", "PAY_AMT6", "default.payment.next.month"))
#Inspect the data frame
library(dplyr)
credit <- credit %>% rename("DEFAULT" = "default.payment.next.month")

```

```{r}
# Missing Values & Duplicates Inspection
sum(is.na(credit))

duplicates <- credit%>%duplicated()
duplicates_amount <- duplicates%>%(table)
duplicates_amount
```

```{r}
# Remove Duplicates
credit <- credit%>%distinct()
# Displays how many duplicates are present in the updated data frame.
duplicates_counts <- credit%>%duplicated()%>%table()
duplicates_counts

```





```{r}
# Restructure DataFrame
credit$SEX[credit$SEX %in% "1"] = "Male"
credit$SEX[credit$SEX %in% "2"] = "Female"

credit$EDUCATION[credit$EDUCATION %in% "0"] = "Unknown"
credit$EDUCATION[credit$EDUCATION %in% "1"] = "Grad School"
credit$EDUCATION[credit$EDUCATION %in% "2"] = "College"
credit$EDUCATION[credit$EDUCATION %in% "3"] = "High School"
credit$EDUCATION[credit$EDUCATION %in% "4"] = "Other"
credit$EDUCATION[credit$EDUCATION %in% "5"] = "Unknown"
credit$EDUCATION[credit$EDUCATION %in% "6"] = "Unknown"

credit$MARRIAGE[credit$MARRIAGE %in% "0"] = "Unknown"
credit$MARRIAGE[credit$MARRIAGE %in% "1"] = "Married"
credit$MARRIAGE[credit$MARRIAGE %in% "2"] = "Single"
credit$MARRIAGE[credit$MARRIAGE %in% "3"] = "Other"

credit$DEFAULT[credit$DEFAULT %in% "0"] = "No"
credit$DEFAULT[credit$DEFAULT %in% "1"] = "Yes"

head(credit)
```

# Data Visualization
# Bar Chart of SEX, EDUCATION, MARRIAGE
```{r}
credit %>%
  ggplot(aes(x = SEX , fill = DEFAULT)) +
  geom_bar()

credit %>%
  ggplot(aes(x = EDUCATION, fill = DEFAULT)) +
  geom_bar()

credit %>%
  ggplot(aes(x = MARRIAGE, fill = DEFAULT)) +
  geom_bar()
```
## Histogram of AGE
```{r}
credit %>%
  ggplot(aes(x = AGE)) + 
  geom_histogram(aes(y = ..density..), fill = "Blue", col = "Grey", binwidth = 5) + 
  geom_density(alpha = 0.2, color = "black", fill = "blue")
```
## Histogram of PAY
```{r}
credit %>%
  ggplot(aes(x = PAY_0)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
credit %>%
  ggplot(aes(x = PAY_2)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
credit %>%
  ggplot(aes(x = PAY_3)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
credit %>%
  ggplot(aes(x = PAY_4)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
credit %>%
  ggplot(aes(x = PAY_5)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
credit %>%
  ggplot(aes(x = PAY_6)) + 
  geom_histogram(aes(fill = DEFAULT), col = "Grey", binwidth = 1)
```

## Boxplots
```{r}
ggplot(credit,aes(x=DEFAULT,y=AGE,fill=DEFAULT))+geom_boxplot()
ggplot(credit,aes(x=DEFAULT,y=AGE,fill=EDUCATION))+geom_boxplot()
ggplot(credit,aes(x=DEFAULT,y=AGE,fill=MARRIAGE))+geom_boxplot()

ggplot(credit,aes(x=SEX,y=LIMIT_BAL,fill=DEFAULT))+geom_boxplot()
ggplot(credit,aes(x=EDUCATION,y=LIMIT_BAL,fill=DEFAULT))+geom_boxplot()
ggplot(credit,aes(x=MARRIAGE,y=LIMIT_BAL,fill=DEFAULT))+geom_boxplot()
```
## Features
```{r}
#Feature 1
#Gold customer is less likely to default
max(credit$LIMIT_BAL)
min(credit$LIMIT_BAL)
mean(credit$LIMIT_BAL)
cut_points <- c(10000, 200000, 500000, 100000)
labels <- c("bronze", "silver", "gold")
credit$cust_cat <- cut(credit$LIMIT_BAL, breaks = cut_points, labels = labels)

credit %>%
  ggplot(aes(x = cust_cat, fill = DEFAULT)) +
  geom_bar()
```


```{r}
#Feature 2
cut_points <- c(21, 40, 60, 81)
labels <- c("young", "middle", "senior")
credit$age_group <- cut(credit$AGE, breaks = cut_points, labels = labels)

credit %>%
  ggplot(aes(x = age_group, fill = DEFAULT)) +
  geom_bar()
```
```{r}
#Feature3
#The new column means the total number of payments done or missed.
credit$pay_total = (credit$PAY_0 + credit$PAY_2 + credit$PAY_3+ credit$PAY_4+ credit$PAY_5+ credit$PAY_6)
mean(credit$pay_total)
#Observations 

#1. Id's with score above -2 shows that the customers have always paid on time.
#2. Id's with positive values shows the number of payment default month by customer. 
#3. Customers have duly paid over a year are good customers.

cut_points <- c(-20, -10, 0, 10)
labels <- c("low", "medium", "high")
credit$risk_cat <- cut(credit$pay_total, breaks = cut_points, labels = labels)

credit %>%
  ggplot(aes(x = risk_cat, fill = DEFAULT)) +
  geom_bar()
```



```{r}
#Feature 4
library(ggExtra)
scatter_plot_age = credit %>%
  ggplot(aes(x = AGE, y = pay_total, color = pay_total)) +
  geom_point() + 
  labs(x = "Age", y = "Pay_total", title = "Scatter Plot") +
  scale_color_gradient(low = "salmon", high = "cyan")
ggMarginal(scatter_plot_age, type = "histogram", fill = "#D2E5D0", col = "#D2E5D0")
#We can clearly see that payment default are being done more by age group of 21 to 55.
```

```{r}
#Feature 5
scatter_plot_education = credit %>%
  ggplot(aes(x = EDUCATION, y = pay_total, color = pay_total)) +
  geom_point() + 
  labs(x = "education", y = "Pay_total", title = "Scatter Plot") +
  scale_color_gradient(low = "salmon", high = "cyan")
ggMarginal(scatter_plot_education, type = "histogram", fill = "#D2E5D0", col = "#D2E5D0")
#We can clearly see that payment default are being done more by 
```
```{r}
# Feature 6
credit %>% ggplot(aes(x = EDUCATION, y = pay_total, fill = risk_cat)) +
  geom_boxplot() +
  labs(x = "Education", y = "Pay_total", title = "Box Plot") +
  theme(legend.position = "top") +
  guides(fill = guide_legend(title = "Risk Category"))

```

```{r}
#Feature 7
credit %>% ggplot(aes(x = age_group, y = pay_total, fill = risk_cat)) +
  geom_violin() +
  labs(x = "Age Group", y = "Pay_total", title = "Violin Plot") +
  theme_minimal() +
  theme(legend.position = "top") +
  guides(fill = guide_legend(title = "Risk Category"))

```
```{r}
# Feature 8
library(ggcorrplot)
part_of_credit = credit[, c('SEX', 'MARRIAGE', 'EDUCATION', 'age_group', 'pay_total', 'risk_cat')]
model.matrix(~0+., data=part_of_credit) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2) +
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5))
```

```{r}
credit = subset(credit, select = c("LIMIT_BAL", "SEX", "EDUCATION", "MARRIAGE", "AGE", "PAY_0",
                                 "PAY_2", "PAY_3", "PAY_4", "PAY_5", "PAY_6",
                                 "BILL_AMT1", "BILL_AMT2", "BILL_AMT3",
                                 "BILL_AMT4", "BILL_AMT5", "BILL_AMT6",
                                 "PAY_AMT1", "PAY_AMT2", "PAY_AMT3","PAY_AMT4",
                                 "PAY_AMT5", "PAY_AMT6", "DEFAULT"))
```



## Resample
```{r}
#Through descriptive results, we found our dataset is imbalanced. we need to oversample by adding more weight to default.
credit_model = credit[sample(1:nrow(credit),30000,prob = ifelse(credit$DEFAULT == 'Yes', 0.8,0.2),replace = TRUE),]
table(credit_model$DEFAULT)
```

## Modeling
```{r}
# Training Datasets & Test Datasets
set.seed(66)
# The data is split with 75% in training and 25% in testing.
sample <- sample(c(TRUE, FALSE), nrow(credit_model), replace = TRUE, prob = c(0.75, 0.25))
train_set <- credit_model[sample, ]
test_set <- credit_model[!sample, ]
```


## Classification Trees
## Classification Trees: Model
```{r}
library(rpart)
library(rpart.plot)
library(caret)
```

```{r}
credit_ct<-rpart(DEFAULT ~ .,  
                data=train_set,                            
                method="class",                           
                control=rpart.control(cp=0,maxdepth = 6))   
```

```{r}
printcp(credit_ct)
```

```{r}
min_xerror_tree<-credit_ct$cptable[which.min(credit_ct$cptable[,"xerror"]),]
# prune tree with minimum cp value
min_xerror_tree<-prune(credit_ct,cp=min_xerror_tree[1])
rpart.plot(min_xerror_tree)
print(min_xerror_tree)
min_xerror_tree$variable.importance
```

```{r}
pred_ct_class<-predict(credit_ct,test_set,type="class") 
head(pred_ct_class)

pred_ct<-predict(credit_ct,test_set)  
head(pred_ct)

pred_ct_prob<-predict(credit_ct,test_set, type = 'prob')[,2]

confusion_matrix_ct <- table(pred_ct_class,test_set$DEFAULT, dnn=c("predicted","actual"))
print(confusion_matrix_ct)
## false positive rate = FP/(FP+TN) = 983/(983+2915) = 28.06%
## false negative rate = FN/(TP+FN) = 1153/(1153+2723) = 29.74%

accuracy_ct <- sum(diag(confusion_matrix_ct))/sum(confusion_matrix_ct)
accuracy_ct
```

## Cross-validation
```{r}
library(caret)
control <- trainControl(method = "cv", number = 10)
# Train the model using cross-validation
credit_cv <- train(DEFAULT ~ ., data = credit_model, method = "rpart", trControl = control)
print(credit_cv$results)
print(credit_cv$finalModel)
rpart.plot(credit_cv$finalModel)

pred_cv_class<-predict(credit_cv,test_set,type="raw") 
head(pred_cv_class)

pred_cv<-predict(credit_cv,test_set)  
head(pred_cv)

pred_cv_prob<- predict(credit_cv,test_set,type = 'prob')[,2]

confusion_matrix_cv <- table(pred_cv_class,test_set$DEFAULT, dnn=c("predicted","actual"))
print(confusion_matrix_cv)
## false positive rate = FP/(FP+TN) = 855/(855+2647) = 24.41%
## false negative rate = FN/(TP+FN) = 1315/(1315+2561) = 33.92%

accuracy_cv <- sum(diag(confusion_matrix_cv))/sum(confusion_matrix_cv)
accuracy_cv
```


## Random Forest
## Random Forest: Model
```{r}
set.seed(66)
library(randomForest)
set.seed(123)
credit_randomforest <- randomForest(factor(DEFAULT) ~., mtry = 5, data = train_set)
credit_randomforest

head(credit_randomforest$votes)    
head(credit_randomforest$predicted) 
```

```{r}
varImpPlot(credit_randomforest)
```

## Random Forest: Prediction
```{r}
pred_randomforest<-predict(credit_randomforest,test_set)
pred_randomforest_prob<- predict(credit_randomforest,test_set,type = 'prob')[,2]
confusion_matrix_randomforest <- table(pred_randomforest,test_set$DEFAULT, dnn=c("predicted","actual"))
print(confusion_matrix_randomforest)
## false positive rate = FP/(FP+TN) = 392/(392+3110) = 11.19%
## false negative rate = FN/(TP+FN) = 259/(259+3617) = 6.68%

accuracy_randomforest <- sum(diag(confusion_matrix_randomforest))/sum(confusion_matrix_randomforest)
accuracy_randomforest
```


## Logistics Regression
## Logistics Regression: Model
```{r}
train_set$DEFAULT = ifelse(train_set$DEFAULT =='Yes',1, 0)
credit_glm <- glm(DEFAULT ~ ., data = train_set, family = binomial())
summary(credit_glm)
```

## Logistics Regression: Prediction
```{r}
pred_glm_prob <- predict.glm(credit_glm, newdata = test_set, type = "response")
head(pred_glm_prob)
pred_glm <- ifelse(pred_glm_prob<0.5,0,1)
confusion_matrix_glm <- table(pred_glm,test_set$DEFAULT,dnn=c("predicted","actual"))
print(confusion_matrix_glm)
## false positive rate = FP/(FP+TN) = 1406/(1406+2096) = 40.14%
## false negative rate = FN/(TP+FN) = 1168/(1168+2708) = 30.13%

accuracy_glm <- sum(diag(confusion_matrix_glm))/sum(confusion_matrix_glm)
accuracy_glm
```


## Support Vector Machine (SVM)
```{r}
library(e1071)
credit_svm <- svm(formula = DEFAULT ~ ., 
               data = train_set,                   
               kernel = "linear",  
               cost = 0.1)       
credit_svm
```

```{r}
dv<-data.frame(credit_svm$decision.values)

ggplot(dv,aes(x=X.)) +
  geom_histogram(colour="black",fill="white")
```

```{r}
pred_svm_prob <- predict(credit_svm,test_set,type="prob")
head(pred_svm_prob)
pred_svm <- ifelse(pred_svm_prob<0.5,0,1)
confusion_matrix_svm <- table(pred_svm, test_set$DEFAULT,dnn=c("predicted","actual"))
print(confusion_matrix_svm)
## false positive rate = FP/(FP+TN) = 1188/(1188+2314) = 34.92%
## false negative rate = FN/(TP+FN) = 1291/(1291+2585) = 33.30%

accuracy_svm <- sum(diag(confusion_matrix_svm))/sum(confusion_matrix_svm)
accuracy_svm
```



```{r}
test_set$DEFAULT_value = ifelse(test_set$DEFAULT == 'Yes', 1,0)

library(pROC)
ct_roc<-roc(test_set$DEFAULT_value,pred_ct_prob,auc=TRUE)
cv_roc<-roc(test_set$DEFAULT_value,pred_cv_prob,auc=TRUE)
randomforest_roc<-roc(test_set$DEFAULT_value,pred_randomforest_prob,auc=TRUE)
logit_roc<-roc(test_set$DEFAULT_value,pred_glm_prob,auc=TRUE)
svm_roc<-roc(test_set$DEFAULT_value,pred_svm_prob,auc=TRUE)

# Plot the ROC curves
plot(ct_roc, print.auc = TRUE, legacy.axes = TRUE, print.auc.y = 0.9, xlab = "False Positive Rate", ylab = "True Positive Rate", col = "#FF771E")
plot(cv_roc,print.auc=TRUE,print.auc.y=.7, col="#00C3F9",add=TRUE, lty = 2)
plot(randomforest_roc, print.auc = TRUE, print.auc.y = 0.5, col = "#995491", add = TRUE, lty = 3)
plot(logit_roc,print.auc=TRUE,print.auc.y=.3, col="#00766C",add=TRUE, lty = 4)
plot(svm_roc,print.auc=TRUE,print.auc.y=.1, col="#BB3E3F",add=TRUE, lty = 5)

legend("bottomright", legend = c("ct", "cv", "rf", "logit", "svm"), col = c("#FF771E", "#00C3F9", "#995491", "#00766C", "#BB3E3F"), lty = c(1, 2, 3, 4, 5))

```
```{r}

```
